---
title: "NBACK AND LURES - Statistical Analysis"
date: 'Thursday,  August 1 08:00 CET'
author: "Deniz Guen"
output: html_document
---

# Required R packages
* `tidyverse` (or `ggplot2`, `dplyr`, `purrr`, `tibble`)
* `brms`

### Importing Libraries

```{r setup, include=FALSE, echo = FALSE, message = FALSE}
library(tidyverse)
library(rstan)
library(brms)
library(lme4)
options(mc.cores = max(parallel::detectCores()-1, 1)) # at least one core if the machine only has one
rstan_options(auto_write = TRUE)
```

### Setting seed for pseudo random operations
```{r setting seed}
set.seed(100) # this ensures that any psuedo-random computations are always the same
```



# Description
This file is a documentation of the statistical procedures that were used to analyse the data from the nback experiment that was deployed for the report 'The Importance of controlling for lure foils in nback tasks' by Deniz M. Gün


### Background: nback task
In the nback task participants are shown a sequence of stimuli. For every stimulus the participants have to indicate wether the same stimulus has been presented exactly n steps earlier. There are different variants that may require participants to pay attention to more than one modality at the same time (e.g. shape and color or number and sound). But for this experiment a simple variant was chosen.

The participants absolved 3- and 4back tasks with 100 trials for each, where the stimuli were integers between 1 and 9. They had 4 seconds to respond for each stimulus. 

There were two groups that differed in the ratio of lure foils to other stimuli. In the control group the lure foils were just as likely as any other number (1/9) while in the main groups the lure foils were assigned with a probability of 1/4 each. The control group is supposed to demonstrate that the lower accuracy for lure foils is not a result of some unknown effect that simply arises from having more lure foils.


### Purpose
The n-back task is sometimes used in comparative studies to assess, e.g. wether a training programm had an effect on performances in working memory related tasks like the nback task. The performances of participants, however, may depend on the proportion of lure foils in the sequence of stimuli. So, this experiment is supposed to demonstrate that different kinds of stimuli which are usually not explicitly marked by experimentors significantly influence the results. Therefore in future experiments other researchers need to compare the performances on specific stimulus types to rule out their influence on their results.   

# 1. Overview

#### Importing data
```{r Import Data,message=TRUE, warning=TRUE, tidy=FALSE}
data.raw = data.frame(read.csv("fake_data.csv"))
head(data.raw)
```

# 2. Data Cleaning

#### Removing non-pertinent data (practice sessions)
This includes every trial number smaller than n. Taking the 4back task for example, a target can not appear before the 4th trial.

```{r DataClean1}
data_pertinent = subset(data.raw, trial_name != "practice" & trial_number > n)
```

#### Removing irrelevant columns
The only relevant columns for now are the stimulus type, the correctness, the reaction time, trial name, n-condition, the trial number and the submission id (subject)

```{r RelevantColumns}
data = subset(data_pertinent, select= c(stim_type,correctness,RT,trial_name,n,trial_number,submission_id))
```

#### Removing obviously Faulty Trials 
Subjects with less than 200 trials overall:
```{r Grouping}
grouped = count(group_by(data, submission_id))
ids_faulty = filter(grouped, n < 200)$submission_id
data = subset(data, !(submission_id %in% ids_faulty))
```

#### Removing outliers with respect to Reaction Times (applying Rosner Test)
Should there be outliers, all entries for a specific submission ID will be removed
```{r Outlier Detection Function}

normalize = function(data) {
  mu = mean(data)
  sig = sd(data)
  normalized = abs(data - mu)/sig
  return(normalized)
}

rosner_outlier_removal = function(data, var.col, crit.value){

  data$normalized = normalize(var.col)
  maximum = c( max(data$normalized), min(data$normalized) ) %>% abs() %>% max()
  
  if(maximum >= crit.value){
    data <- subset(data, abs(data$normalized) != maximum)
    return(rosner_test(data, data$normalized, crit.value))
    
  } else {
    return(data)
  }
}

get_outliers = function(data, var.col,grouping.variables, crit.value){
  #woo : "without outliers"
  data.woo = rosner_outlier_removal(data=data, var.col=var.col, crit.value=crit.value)
  data.woo = select(data.woo, c(1:(length(names(data.woo)))))
  
  all = select(data, grouping.variables)
  non_outliers = select(data.woo, grouping.variables)
  outliers = anti_join(all,non_outliers)
  return(outliers)
}


```

We find the outliers based on unlikely reaction times using the rosner test for outlier detection ( and removal).
```{r Finding Outliers, eval=FALSE, include=FALSE}
mean_rts = aggregate(data$RT, FUN=mean, by=list(data$submission_id, data$n))
names(mean_rts) = c("submission_id", "n", "x")

outliers = get_outliers(data=mean_rts, var.col=mean_rts$x, 
                         grouping.variables=c("submission_id", "n"), crit.value=3)
outliers
```

Now we remove the outliers from our dataset.
```{r Removing outliers from data, eval=FALSE, include=FALSE}
if(nrow(outliers)>0){
  for(row in c(1:nrow(outliers))){
    sprintf("ID: %d, n: %d",outliers$submission_id[row], outliers$n[row] )
    data <- subset(data, submission_id != outliers$submission_id[row] , n != outliers$n[row])
  }
}
```


#### Adding column stim_type_gross
Could be interesting for plotting.
```{r Add gross stimtype}
gross = data$stim_type
gross = ifelse(data$stim_type == "target", "target", "non-target")
data$gross_stimtype = gross
```


#### Creating a table with accuracy for different IDs and stimuli

The grouping is important for counting the stimuli for a specific combination of values (e.g.: ID = 1, stim_type = "non-target", correctness = "correct", n=3, etc..)  such that the acuracy for a specific kind of stimulus can be calculated.

```{r count group members}
#Grouping by ID stim type, trial name and n --- 'nn' is the total number of trials for the specific variable combination of each row
grouped1 = data %>% group_by(submission_id,stim_type,gross_stimtype,trial_name,n)
cg1 = count(grouped1)
#Grouping by correctness for counts of "correct" or "incorrect" --- 'nn' is number of correct trials for the specific variable combination of each row
grouped2 = grouped1 %>% group_by(correctness, add=TRUE)
cg2 = count(grouped2)
# Getting list with unique values to iterate through in the following cells
IDs = unique(grouped1$submission_id)
stim_types = unique(grouped1$stim_type)
gross_stimtypes = unique(grouped1$gross_stimtype)
trial_names = c("main", "control")
N = c("3", "4")
```

##### Creating an accuracy table template

```{r Accuracy Table template}
# Accuracy table must be a table with the following rows: ID, trial_name, n, stimulus_type, gross_stimulus_type and accuracy
accuracy_table <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(accuracy_table) <- c("ID","trial_name", "stimulus_type","gross_stimulus_type","n", "accuracy")

# Defining the right variable types for cols.
accuracy_table$submission_id <- as.factor(accuracy_table$submission_id)
accuracy_table$accuracy <- as.numeric(accuracy_table$accuracy)
accuracy_table$n <- as.factor(accuracy_table$n)

```

###### Fill a complete accuracy table ordered on all relevant variables.

```{r Fill a complete accu table}
accu_complete <- subset(accuracy_table, select=c(ID, trial_name, stimulus_type, gross_stimulus_type, n, accuracy))

for(id in IDs){
  for(n in N){
    for(st in stim_types){
    
      howmanyback = n
      trial_name = filter(data,submission_id==id)$trial_name[1]
      gr_st = filter(data, stim_type == st)$gross_stimtype[1]
      n_correct = sum( filter(cg2, (submission_id == id && stim_type == st) && (n == howmanyback && correctness == "correct"))$nn )
      n_total = sum( filter(cg2, (submission_id == id && n == howmanyback) && stim_type == st)$nn )
      accuracy = n_correct/n_total
      
      entry = data.frame(submission_id = id, trial_name = trial_name, stimulus_type = st, gross_stimulus_type = gr_st, n = howmanyback, accuracy = accuracy, count = n_total)
      accu_complete <- rbind(accu_complete, entry)
    }
  }
}
head(accu_complete)
```

##### Variables
Dependent: Accuracy
Independent: trial_name, (gross_)stimulus_type, id, n, count

```{r Setting Variable Types}
accu_complete$submission_id <- as.factor(accu_complete$submission_id)
accu_complete$n <- as.factor(accu_complete$n)
```


# 3. Hypothesis Tests

#### The proposed hypotheses were

$H1_a$ and $H1_b$: The accuracy for lure foils at n-1 ($a$) as well as at n+1($b$)) is lower than the accuracy of non-targets. 

$H0_c$: The accuracy of lure foils at n+1 does not differ significantly from lure foils at n-1.

###### Getting an Overview

First we get an overview to illustrate the data
```{r Boxplot Accuracy}
ggplot(accu_complete,mapping=aes(x=gross_stimulus_type, y=accuracy, color=n))+
  geom_boxplot()
```


Now we remove outliers with respect to accuracy and plot again.

```{r Removing Outliers and plotting again, eval=FALSE, include=FALSE}
accu_complete <- rosner_outlier_removal(accu_complete, var.col=accu_complete$accuracy, crit.value=3)
ggplot(data=accu_complete.woo, mapping=aes(x=stimulus_type, y=accuracy, color=n))+
  geom_boxplot()

```

### Model
```{r Listing the Variables}
names(accu_complete)
```

We'll make accuracy in our formula dependend on the fixed effects stimulus_type and n and include the random intercepts for subjects (="submission_id") and groups (="trial_name")

```{r Regression Formula}
formula = bf(accuracy ~ stimulus_type*n+(1|submission_id)+(1|trial_name),
          sigma ~ stimulus_type*n+(1|submission_id) + (1|trial_name))
```

Now we create a bayesian regression model where we assume student's distribution (Kruschker's Model: Robust Bayesian Estimation)
```{r Creating brms model }
accu_complete.mdl.brm = brm(formula, data=accu_complete, family="student")
```


# Testing Hypotheses
In order to test our hypotheses we look at the coefficients from our posterior samples, which we obtained from the model above using MCMC. 

If a coefficient, say e.g. "b_stimulus_typetarget" of one sample is smaller than 0, this indicates the following:
Based on our estimates for the real parameters, there is a sample drawn from our posterior distribution for which it is true, that the mean accuracy for the stimulus type "target" is lower than the mean accuracy for the stimulus type "non-target" which is our intercept. 

What we are going to do in order to assess the probabilities of our hypotheses, is, look at the ratio of samples for which a coefficient like e.g. "b_stimulus_typetarget" is smaller than 0.


We predict that the accuracy is lower for all kinds of lures than it is for non-targets. We'll accept in a preliminary fashion that our hypotheses is true if the probability that one of the mentioned coefficients is lower than the intercept is at least 95%. 


```{r}
accu_complete.mdl.brm
```

We can't always tell beforehand which value is chosen as intercept. Therefore this has to be done again after all the data has been collected. But in this case "non-target" seems to be our intercept.
$H1:$ _Lure foils are less probable than non-targets._
```{r Calculating probabilities of Hypotheses}
post.samples = posterior_samples(accu_complete.mdl.brm)
#For n3
p.H1a = mean(post.samples$b_stimulus_typelureMnM1 < post.samples$b_Intercept)
p.H1b = mean(post.samples$b_stimulus_typelureMnP1 < post.samples$b_Intercept)

p.hypotheses = data.frame(H1a = p.H1a, H1b = p.H1b)

#For the interaction with n4
p.H1a.n4 = mean(post.samples$b_stimulus_typelureMnM1 +
               post.samples$"b_stimulus_typelureMnM1:n4" +
               post.samples$"b_n4" < post.samples$b_Intercept)

p.H1b.n4 = mean(post.samples$b_stimulus_typelureMnP1 +
               post.samples$"b_stimulus_typelureMnP1:n4" +
               post.samples$"b_n4" < post.samples$b_Intercept)

p.hypotheses = rbind(p.hypotheses, data.frame(H1a = p.H1a.n4, H1b = p.H1b.n4))
p.hypotheses$n <- as.factor(c(3,4))

print(p.hypotheses)
```
For our fake data none of the samples yielded a mean for the coefficients of interests that is higher than the mean for the parameter non-target. Therefore it is likely that both alternative hypotheses $H1_a$ and $H1_b$ should be accepted and $H0$ should be rejected.



$H1c :$  _Lure foils at n+1 are significantly more accurately classified than lures at n-1._

We use a Kolgomorov-Smirnov Test. We assess the sameness of the parameters "stimulus_typelure-n+1" and "stimulus_type_lure-n-1". The p-values are based on a statistic that measures the maximum Distance between the empirical cumulative distribution functions of both parameters. High p-values indicate the probability to see a bigger difference if those two samples are from the same distribution. A very low p-value indicates that there is not a good statistical support for the difference between the two parameters.

```{r}

# Testing wether lures at n+1 are significantly more accurate than lures at n-1
#[2] Is the p.value of the summary that is returned by ks()
p.H1c = ks.test(post.samples$b_stimulus_typelureMnP1, post.samples$b_stimulus_typelureMnM1)[2]
#Considering the interaction with and the intercept for n=4
p.H1c.n4 = ks.test(post.samples$b_stimulus_typelureMnP1 +
                  post.samples$"b_stimulus_typelureMnP1:n4"
                , post.samples$b_stimulus_typelureMnM1 +
                  post.samples$"b_stimulus_typelureMnM1:n4")[2]

p.hypotheses <- add_column(p.hypotheses, c(p.H1c, p.H1c.n4), .before="n")
colnames(p.hypotheses)[3] <- "H1c"
print(p.hypotheses)
```




# Code Summary
```{r echo=TRUE, eval=FALSE}
library(tidyverse)
library(rstan)
library(brms)
library(lme4)
options(mc.cores = max(parallel::detectCores()-1, 1)) # at least one core if the machine only has one
rstan_options(auto_write = TRUE)
set.seed(100)

data.raw = data.frame(read.csv("fake_data.csv"))
data_pertinent = subset(data.raw, trial_name != "practice" & trial_number > n)
data = subset(data_pertinent, select= c(stim_type,correctness,RT,trial_name,n,trial_number,submission_id))

# Removing obviously Faulty Trials 
grouped = count(group_by(data, submission_id))
ids_faulty = filter(grouped, n < 200)$submission_id
data = subset(data, !(submission_id %in% ids_faulty))

#Defining Functions for outlier detection and removal
#####################################################################################
normalize = function(data) {
  mu = mean(data)
  sig = sd(data)
  normalized = abs(data - mu)/sig
  return(normalized)
}

rosner_outlier_removal = function(data, var.col, crit.value){

  data$normalized = normalize(var.col)
  maximum = c( max(data$normalized), min(data$normalized) ) %>% abs() %>% max()
  
  if(maximum >= crit.value){
    data <- subset(data, abs(data$normalized) != maximum)
    return(rosner_test(data, data$normalized, crit.value))
    
  } else {
    return(data)
  }
}

get_outliers = function(data, var.col,grouping.variables, crit.value){
  #woo : "without outliers"
  data.woo = rosner_outlier_removal(data=data, var.col=var.col, crit.value=crit.value)
  data.woo = select(data.woo, c(1:(length(names(data.woo)))))
  
  all = select(data, grouping.variables)
  non_outliers = select(data.woo, grouping.variables)
  outliers = anti_join(all,non_outliers)
  return(outliers)
}
##############################################################################################################################
mean_rts = aggregate(data$RT, FUN=mean, by=list(data$submission_id, data$n))
names(mean_rts) = c("submission_id", "n", "x")

outliers = get_outliers(data=mean_rts, var.col=mean_rts$x, 
                         grouping.variables=c("submission_id", "n"), crit.value=3)
if(nrow(outliers)>0){
  for(row in c(1:nrow(outliers))){
    sprintf("ID: %d, n: %d",outliers$submission_id[row], outliers$n[row] )
    data <- subset(data, submission_id != outliers$submission_id[row] , n != outliers$n[row])
  }
}

# Adding column stim_type_gross
gross = data$stim_type
gross = ifelse(data$stim_type == "target", "target", "non-target")
data$gross_stimtype = gross

#Grouping by ID stim type, trial name and n --- 'nn' is the total number of trials for the specific variable combination of each row
grouped1 = data %>% group_by(submission_id,stim_type,gross_stimtype,trial_name,n)
cg1 = count(grouped1)
#Grouping by correctness for counts of "correct" or "incorrect" --- 'nn' is number of correct trials for the specific variable combination of each row
grouped2 = grouped1 %>% group_by(correctness, add=TRUE)
cg2 = count(grouped2)
# Getting list with unique values to iterate through in the following cells
IDs = unique(grouped1$submission_id)
stim_types = unique(grouped1$stim_type)
gross_stimtypes = unique(grouped1$gross_stimtype)
trial_names = c("main", "control")
N = c("3", "4")

# Accuracy table must be a table with the following rows: ID, trial_name, n, stimulus_type, gross_stimulus_type and accuracy
accuracy_table <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(accuracy_table) <- c("ID","trial_name", "stimulus_type","gross_stimulus_type","n", "accuracy")

# Defining the right variable types for cols.
accuracy_table$submission_id <- as.factor(accuracy_table$submission_id)
accuracy_table$accuracy <- as.numeric(accuracy_table$accuracy)
accuracy_table$n <- as.factor(accuracy_table$n)
accu_complete <- subset(accuracy_table, select=c(ID, trial_name, stimulus_type, gross_stimulus_type, n, accuracy))

for(id in IDs){
  for(n in N){
    for(st in stim_types){
    
      howmanyback = n
      trial_name = filter(data,submission_id==id)$trial_name[1]
      gr_st = filter(data, stim_type == st)$gross_stimtype[1]
      n_correct = sum( filter(cg2, (submission_id == id && stim_type == st) && (n == howmanyback && correctness == "correct"))$nn )
      n_total = sum( filter(cg2, (submission_id == id && n == howmanyback) && stim_type == st)$nn )
      accuracy = n_correct/n_total
      
      entry = data.frame(submission_id = id, trial_name = trial_name, stimulus_type = st, gross_stimulus_type = gr_st, n = howmanyback, accuracy = accuracy, count = n_total)
      accu_complete <- rbind(accu_complete, entry)
    }
  }
}
accu_complete$submission_id <- as.factor(accu_complete$submission_id)
accu_complete$n <- as.factor(accu_complete$n)

#Plotting Data
ggplot(accu_complete,mapping=aes(x=gross_stimulus_type, y=accuracy, color=n))+
  geom_boxplot()

accu_complete <- rosner_outlier_removal(accu_complete, var.col=accu_complete$accuracy, crit.value=3)
ggplot(data=accu_complete, mapping=aes(x=stimulus_type, y=accuracy, color=n))+
  geom_boxplot()
names(accu_complete)

# Creating the Model
formula = bf(accuracy ~ stimulus_type*n+(1|submission_id)+(1|trial_name),
          sigma ~ stimulus_type*n+(1|submission_id) + (1|trial_name))
accu_complete.mdl.brm = brm(formula, data=accu_complete, family="student")
#Testing Hypotheses
post.samples = posterior_samples(accu_complete.mdl.brm)
p.H1a = mean(post.samples$b_stimulus_typelureMnM1 < post.samples$b_stimulus_typenonMtarget)
p.H1b = mean(post.samples$b_stimulus_typelureMnP1 < post.samples$b_stimulus_typenonMtarget)
p.hypotheses = data.frame(H1a = p.H1a, H1b = p.H1b)
# Interaction with n=4
p.H1a.n4 = mean(post.samples$b_stimulus_typelureMnM1 +
               post.samples$"b_stimulus_typelureMnM1:n4" +
               post.samples$"b_n4" < post.samples$b_stimulus_typenonMtarget)
p.H1b.n4 = mean(post.samples$b_stimulus_typelureMnP1 +
               post.samples$"b_stimulus_typelureMnP1:n4" +
               post.samples$"b_n4" < post.samples$b_stimulus_typenonMtarget)
p.hypotheses = rbind(p.hypotheses, data.frame(H1a = p.H1a.n4, H1b = p.H1b.n4))
p.hypotheses$n <- as.factor(c(3,4))
print(p.hypotheses)
p.H1c = ks.test(post.samples$b_stimulus_typelureMnP1, post.samples$b_stimulus_typelureMnM1)[2]
#Considering the interaction with and the intercept for n=4
p.H1c.n4 = ks.test(post.samples$b_stimulus_typelureMnP1 +
                  post.samples$"b_stimulus_typelureMnP1:n4"
                , post.samples$b_stimulus_typelureMnM1 +
                  post.samples$"b_stimulus_typelureMnM1:n4")[2]
p.hypotheses <- add_column(p.hypotheses, c(p.H1c, p.H1c.n4), .before="n")
colnames(p.hypotheses)[3] <- "H1c"
print(p.hypotheses)
```


